
```{css, echo=FALSE}
body{
  background-color: #FAFAFA;
    font-size: 18px;
  line-height: 1.8;
}
code.r{
  font-size: 16px;
}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this project, we are interested in predicting the direction in the foreign exchange market in Turkey. To this, we will train a model to classify the movements in the USD/TL exchange rate. The data set covers the period 13/01/2000 - 04/05/2020 and contains 5086 daily observations.

usdret is today’s return in the USD/TL exchange rate and it is defined as the daily percentage change in the USD/TL exchange rate. usddirection is simply the sign of usdret and it has two levels: “Down” if usdret is negative, and “Up” if usdret is positive.

The data set also contains daily percentage returns on the BIST100 stock index, bistret and its lags bistretlag1-bistretlag5.
```{r message=FALSE, warning=FALSE}
library(tidyverse)  
load("finmarkets.RData")
str(finmarkets)
head(finmarkets)
``` 

```{r message=FALSE, warning=FALSE}
contrasts(finmarkets$usddirection)
#frequencies
table(finmarkets$usddirection)/nrow(finmarkets)
#the market went down about 50% of the time in the whole sample.
```


```{r}
#Splitting data into train and test set.
finmarkets.train <- finmarkets %>% dplyr::filter(year<=2016)
finmarkets.test <- finmarkets %>% dplyr::filter(year>2016)
```
Data is prepared. Now we'll fit logistic regression.

```{r}
#Logistic Regression
#our predictors are usdretlag1 + usdretlag2 + usdretlag3+usdretlag4 + usdretlag5.
glm.fit <- glm(usddirection ~ usdretlag1 + usdretlag2 + usdretlag3 + 
               usdretlag4 + usdretlag5, data=finmarkets.train, 
               family=binomial)
summary(glm.fit)
broom::tidy(glm.fit)
```


```{r}
#Results from the logistic regression suggests that only the first two lags are statistically significant.(p-value)
#Marginal Effects
library(margins)
ame <- margins::margins(glm.fit)
summary(ame)
plot(ame)

```

If there is one percentage point increase in yesterday’s USD/TL returns, today’s probability of going up will increase by 0.025.
```{r}
#Confusion Matrix
library(caret)
library(e1071)
glm.probs.test <- predict(glm.fit, finmarkets.test, type = "response") 
glm.pred.test  <- rep("Down", nrow(finmarkets.test))
glm.pred.test[glm.probs.test>0.5] = "Up"
glm.pred.test  <- as.factor(glm.pred.test)
confusionMatrix(glm.pred.test, finmarkets.test$usddirection, positive = "Up")
```

Accuracy Rate is higher than "No Information Rate". But difference is not statistically significant.
 P-Value [Acc > NIR] : 0.3515.
So this model above is useless for now.
```{r}

#Let's try to make the model better by using data from the Istanbul stock market.
#new predictors: bistretlag1 + bistretlag2 + bistretlag3 + bistretlag4 + bistretlag5.

glm.fit.stock <- glm(usddirection ~ usdretlag1 + usdretlag2 + usdretlag3 + 
                 usdretlag4 + usdretlag5 + bistretlag1 + bistretlag2 +
                 bistretlag3 + bistretlag4 + bistretlag5, 
                 data=finmarkets.train, family=binomial)
summary(glm.fit.stock)

```
Results showing that we may have some new useful predictors :). bistretlag1, bistretlag2.And we already have usdretlag2.

```{r}
#Let's check accuracy rate with training set
glm.probs.stock <- predict(glm.fit.stock, type = "response") 
glm.pred.stock  <- rep("Down", nrow(finmarkets.train))
glm.pred.stock[glm.probs.stock > 0.5] = "Up"
glm.pred.stock  <- as.factor(glm.pred.stock)
# both predictions and actual must be factors with the same levels
confusionMatrix(glm.pred.stock, finmarkets.train$usddirection, positive = "Up")
```
#Accuracy Rate is 64%! It is also better than no information rate statistically.
#P-Value [Acc > NIR] : < 2.2e-16 .

```{r}
#Now lets observe accuracy rate on test data.
glm.probs.stock.test <- predict(glm.fit.stock, finmarkets.test, type = "response") 
glm.pred.stock.test  <- rep("Down", nrow(finmarkets.test))
glm.pred.stock.test[glm.probs.stock.test>0.5] = "Up"
glm.pred.stock.test  <- as.factor(glm.pred.stock.test)

confusionMatrix(glm.pred.stock.test, finmarkets.test$usddirection, positive = "Up")
```

Accuracy Rate 61.1% is statistically significant when we compare it with NIR. We can say that BIST predictors have made our model much more successful.







----------------------------------------------------------------

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
